HO unification from object language to meta language

### 2024-04-04 paper/main.tex
add paragraph for beta approximation

\subsection{Heuristics}

Possiamo perdere OC e rimanere sospesi o commit di soluzione approx
\begin{elpicode}
  triple ok (@lam x\ @app[@f, @app[@X, x]]) @Y,
  triple ok @X (@lam x\ x),
  triple ok @Y @f
\end{elpicode}

\begin{elpicode}
% @okl 22 [
%   triple ok (@lam x\ @lam y\ @app[@Y, y, x]) @X,
%   triple ok (@lam x\ @f) @X,
% ].
\end{elpicode}

% \section{First order approximation}

% \otext{Coq can solve this: \coqIn{f 1 2 = X 2}, by setting X to f 1}
% \otext{We can re-use part of the algo for $\beta$ given before}

% \section{Unif encoding in real life}
% \otext{Il ML presentato qui è esattamente elpi}
% \otext{Il OL presentato qui è esattamente coq}
% \otext{Come implementatiamo tutto ciò nel solver}

% \section{Results: stdpp and tlc}
% \otext{How may rule are we solving?}
% \otext{Can we do some perf test}

>>>>

\subsection{Relaxing \cref{lemma:keep-fidelity}}

\newcommand{\progBetaNoLLWait}{\emph{progress-beta-\maybebeta}}

% from https://www.cs.mcgill.ca/~bpientka/papers/unif_miller60.pdf
Working with terms in \llambda is sometime too restrictive. There exists systems
such as $\lambda$Prolog~\cite{lamProlog}, Abella~\cite{gacek2008abella}, which 
delay the resolution of \maybebeta unification problems if the substitution
is not able to put them in \llambda.

\printAlll
  {{{(X\appsep a),a},
    {X,\lambda x.Y}}}
  {{}}
  {{}}
  {{}}

In the example above, $\foUnifPb_1$ is in \maybebeta and the object language
cannot solve it, and, by \cref{prop:fidelity}, the meta language neither.
However, we can be more permissive, and relax \cref{lemma:keep-fidelity}. This
modification is quite simple to manage: we are introducing a new \maybebeta
progress rule, say \progBetaNoLLWait, by which, if \lhs is rigid and \rhs is
flexible, the considered \linkbeta is kept in the store and no progression is
done\footnote{This new rule trivially guarantees the termination of progress}.
\progBetaNoLLWait makes \elpiIn{occur-check-links} partial, since the check is
possible only on links with a variable on the \lhs. This means that we can have
two links \linkbetaM{}{X}{Y\appsep a} and \linkbetaM{}{f\appsep X}{Y\appsep a}
where the occur check does not thow an error. Note however, that the
decompilation of the two links will force the unification of $X$ to $Y\appsep a$
and then the unification of $f\appsep (Y \appsep a)$ to $Y\appsep a$, which
fails by the occur check of \Ue.

A second strategy to deal with problem that are in \maybebeta is to make some
approximation. This is the case for example of the unification algorithm of Coq
used in its type class solver~\cite{sozeau08}. The approximation consists in
forcing a choice (among the others) when the unification problem is in
\maybebeta. For instance, in $X \appsep a \appsep b = Y \appsep b$, the last
argument of the two terms is the same, therefore $Y$ is assigned to $X a$. Note
that this is of course an approximation, since $\sigma = \{X = \lambda x.Y, Y =
\_\}$ is another valid substitution for the original problem. This approximation
can be easily introduced in our unification procedure, by adding new custom 
\linkbeta progress rules.
